---
title: "Pyramide des ages"
author: "jcb"
date: "22 août 2015"
output:
  html_document:
    number_sections: yes
    toc: yes
---

Solution avec epicalc
=====================

Todo

Solution avec 'pyramid'
======================

Application: http://freakonometrics.hypotheses.org/17885#more-17885

Drawing population pyramid using data.frame
--------------------------------------------
Usage
```{r pyramid, echo=FALSE}
library(pyramid)
```

 pyramid(data, Laxis=NULL, Raxis=NULL, 
 AxisFM="g", AxisBM="", AxisBI=3, Cgap=0.3, Cstep=1, Csize=1, 
 Llab="Males", Rlab="Females", Clab="Ages", GL=TRUE, Cadj=-0.03, 
 Lcol="Cyan", Rcol="Pink", Ldens=-1, Rdens=-1, main="", ...)
Arguments

__data__    A data.frame including left pyramid numbers in the 1st column and and right pyramid numbers in the 2nd column, where the numbers of males in each age-class are usually given to left numbers and those of females are to right numbers. If the data.frame includes 3rd column, it is used as age-class labels, otherwise the row.names(data) is used as age-class labels.

Laxis	A vector of axis for left pyramid. If missing, automatically given.

Raxis	A vector of axis for right pyramid. If missing, Laxis is used.

AxisFM	A format code of formatC for plotting axis. If missing, "g" is used.

AxisBM	A big.mark of formatC for plotting axis. If missing, none.

AxisBI	A big.interval number of formatC for plotting axis. Default is 3

Cgap	The width of center gap (as ratio to each panel) to draw age-class. Default is 0.3

Cstep	The interval to write the labels of age classes. Default is 1

Csize	The font size factor to write the labels of age classes. Default is 1

Cadj	The vertical adjustment factor for the labels of age classes. Default is -0.03

Llab	The label of the left pyramid. Default is "Males".

Rlab	The label of the right pyramid. Default is "Females".

Clab	The label of the center age-class. Default is "Ages".

GL	Logical value to draw the vertical dotted lines. Default is TRUE.

Lcol    The color of the left pyramid. Default is "Cyan".

Ldens	The density of hatching lines (/inch) for left pyramid. Default is -1, when the pyramid will be filled.

Rcol	The color of the right pyramid. Default is "Pink".

Rdens	The density of hatching lines (/inch) for right pyramid. Default is -1, when the pyramid will be filled.

main	The main title of the pyramid.

...	Other options.

Examples:

```{r}
ages <- c('0-9','10-19','20-29','30-39','40-49','50-59','60-')
males <- c(34,19,11,11,8,7,5)
females <- c(26,25,16,11,7,5,1)
data <- data.frame(males,females,ages)
pyramid(data)

# another example
py.Males <- c(80,40,30,20,10)
names(py.Males) <- c('0-9','10-19','20-29','30-39','40-')
py.Females <- c(60,50,40,30,5)
names(py.Females) <- names(py.Males)
py.df <- data.frame(py.Females,py.Males)
pyramid(py.df,Llab="Females",Rlab="Males",Lcol="navy", Ldens=5, Rcol="red", 
 Rdens=10, GL=FALSE,main="An example of population pyramid\n with auto-axis")

# GunmaPop2005 is included in this package.
pyramid(GunmaPop2005,Llab="Males",Rlab="Females",Clab="",Laxis=seq(0,20000,len=5),
 AxisFM="d", AxisBM=",", Csize=0.8,  Cstep=10, 
 main="Population pyramid of Gunma Prefecture\n (Data: Census 2005, total by gender)")

```

pyramidf
--------

Drawing population pyramid using data.frame within the specified frame

Usage

 pyramidf(data, Laxis=NULL, Raxis=NULL, 
 frame=c(-1.15, 1.15, -0.05, 1.1), 
 AxisFM="g", AxisBM="", AxisBI=3, Cgap=0.3, Cstep=1, Csize=1, 
 Llab="Males", Rlab="Females", Clab="Ages", GL=TRUE, Cadj=-0.03, 
 Lcol="Cyan", Rcol="Pink", Ldens=-1, Rdens=-1, main="", ...)
Arguments

data    
A data.frame including left pyramid numbers in the 1st column and and right pyramid numbers in the 2nd column, where the numbers of males in each age-class are usually given to left numbers and those of females are to right numbers. If the data.frame includes 3rd column, it is used as age-class labels, otherwise the row.names(data) is used as age-class labels.

Laxis	
A vector of axis for left pyramid. If missing, automatically given.

Raxis	
A vector of axis for right pyramid. If missing, Laxis is used.

frame	
A vector of frame c(x-left, x-right, y-bottom, y-top), to draw pyramid. Default frame is same as

AxisFM	
A format code of formatC for plotting axis. If missing, "g" is used.

AxisBM	
A big.mark of formatC for plotting axis. If missing, none.

AxisBI	
A big.interval number of formatC for plotting axis. Default is 3

Cgap	
The width of center gap (as ratio to each panel) to draw age-class. Default is 0.3

Cstep	
The interval to write the labels of age classes. Default is 1

Csize	
The font size factor to write the labels of age classes. Default is 1

Cadj	
The vertical adjustment factor for the labels of age classes. Default is -0.03

Llab	
The label of the left pyramid. Default is "Males".

Rlab	
The label of the right pyramid. Default is "Females".

Clab	
The label of the center age-class. Default is "Ages".

GL	
Logical value to draw the vertical dotted lines. Default is TRUE.

Lcol	
The color of the left pyramid. Default is "Cyan".

Ldens	
The density of hatching lines (/inch) for left pyramid. Default is -1, when the pyramid will be filled.

Rcol	
The color of the right pyramid. Default is "Pink".

Rdens	
The density of hatching lines (/inch) for right pyramid. Default is -1, when the pyramid will be filled.

main	
The main title of the pyramid.

...	
Other options.

Author(s)

Minato Nakazawa minato-nakazawa@umin.net http://minato.sip21c.org/

Examples

```{r}

# GunmaPop2005 and GumaPop2010 are included in this package.
plot(c(0,100), c(0,100), type="n", frame=FALSE, axes=FALSE, xlab="", ylab="", 
 main="Population pyramid of Gunma prefecture")
# overlay
pyramidf(GunmaPop2005, frame=c(10, 75, 0, 90), Clab="", Lcol="skyblue", Rcol="pink",
 Cstep=10, Laxis=0:4*5000, AxisFM="d")
pyramidf(GunmaPop2010, frame=c(10, 75, 0, 90), Clab="", Lcol="deepskyblue", Rcol="deeppink",
 Ldens=10, Rdens=10, Cstep=10, Laxis=0:4*5000, AxisFM="d")
legend("right", fill=c("skyblue","pink","deepskyblue","deeppink"), density=c(NA,NA,10,10), 
 legend=c("Male 2005", "Female 2005", "Male 2010", "Female 2010"), cex=0.8)
```

pyramids
--------
Drawing population pyramid using vectors

Description

Drawing population pyramid using vectors. Detailed explanation is given in Japanese at http://minato.sip21c.org/swtips/R.html#PYRAMID.

Usage

pyramids(Left, Right, Center=NULL, ...)
Arguments

Left    
A vector of the numbers of people given for the left pyramid.

Right	
A vector of the numbers of people given for the right pyramid.

Center	
If given, a vector of the age-class labels to be drawn in center. If not given, names(Left) is used instead.

...	
This is the lapper function for pyramid(). All options for pyramid() can be passed.

Author(s)

Minato Nakazawa minato-nakazawa@umin.net http://minato.sip21c.org/

Examples
```{r}
py.Males <- c(80,40,30,20,10)
 names(py.Males) <- c('0-9','10-19','20-29','30-39','40-')
 py.Females <- c(60,50,40,30,5)
 pyramids(Left=py.Males,Llab="Males",Right=py.Females,Rlab="Females",
  Laxis=c(0,50,100),main="An example of population pyramid\n with fixed axis")
```


Solution avec rCharts
=====================

- Article: Interactive international population pyramids with R, rCharts
- Auteur: Sat 07 June 2014 by Kyle Walker
- source: http://walkerke.github.io/2014/06/rcharts-pyramids/
- source GitHub: http://walkerke.github.io/2014/06/rcharts-pyramids/

Donne plusieurs solutions pour plusieurs bibbliothèques javascript incluses dans __rCharts__. Toutes font appel à la fonction _getAgeTable_ pour récupérer les données. C'est cette fonction qu'il faut adapter pour les données de l'INSEE par exemple. Ce programme permet de faire défiler plsieurs années.

Code
----
## Script to generate population pyramids from the Census Bureau's International Database with rCharts ##
```{}

library(XML)
library(reshape2)
library(rCharts)
library(plyr)


getAgeTable <- function(country, year) {
  c1 <- "http://www.census.gov/population/international/data/idb/region.php?N=%20Results%20&T=10&A=separate&RT=0&Y="  
  c2 <- "&R=-1&C="
  yrs <- gsub(" ", "", toString(year))
  url <- paste0(c1, yrs, c2, country)
  df <- data.frame(readHTMLTable(url))
  nms <- c("Year", "Age", "total", "Male", "Female", "percent", "pctMale", "pctFemale", "sexratio")  
  names(df) <- nms  
  cols <- c(1, 3:9)
  df[,cols] <- apply(df[,cols], 2, function(x) as.numeric(as.character(gsub(",", "", x))))
  df <- df[df$Age != 'Total', ]  
  ord <- 1:nrow(df)
  df <- cbind(df, ord)
  return(df)
}


# DimpleJS pyramid

dPyramid <- function(country, year, colors=NULL) {
  dat <- getAgeTable(country, year)
  dat$Male <- -1 * dat$Male
  
  keep <- c("Year", "Age", "Male", "Female", "ord")
  
  dat.sub <- dat[,keep]
  
  dat.melt <- melt(dat.sub, 
                   value.name='Population', 
                   variable.name = 'Gender', 
                   id.vars=c('Age', 'ord', 'Year') )
  
  dat.melt$gencode <- ifelse(dat.melt$Gender == 'Male', 1, 2)
  
  d1 <- dPlot(
    x = "Population", 
    y = "Age", 
    groups = "Gender", 
    data = dat.melt, 
    type = 'bar')
  
  
  d1$yAxis(type = "addCategoryAxis", orderRule = "ord")
  d1$xAxis(type = "addMeasureAxis")
  d1$legend( x = 60, y = 10, width = 700, height = 20, horizontalAlign = "right")
  
  if (!is.null(colors)){
    d1$colorAxis(
      type = "addColorAxis", 
      colorSeries = "gencode", 
      palette = colors
    )
  }
  if (length(year) > 1) {
    d1$set(storyboard = "Year")
    max_x <- round_any(max(dat.melt$Population), 10000, f = ceiling)
    min_x <- round_any(min(dat.melt$Population), 10000, f = floor)
    d1$xAxis(overrideMax = max_x, overrideMin = min_x)
  }
  
  if (max(dat.melt$Population >= 1000000)) {
    d1$setTemplate( afterScript = 
                      "
                  <script>
                    x._getFormat = function () {
                    return function(d) {
                    return d3.format(',.1f')(Math.abs(d) / 1000000) + 'm';
                     };
                    };
                  myChart.draw()
                  </script>
                  ")
  } else {
    d1$setTemplate( afterScript = 
                      "
                  <script>
                    x._getFormat = function () {
                    return function(d) {
                    return d3.format(',.0f')(Math.abs(d) / 1000) + 'k';
                     };
                    };
                  myChart.draw()
                  </script>
                  ")
  }

  d1
}

# Highcharts pyramid

hPyramid <- function(country, year, colors = NULL) {
  dat <- getAgeTable(country, year)
  dat$Male <- -1 * dat$Male
  
  dat$Age <- factor(dat$Age, levels = rev(dat$Age), labels = rev(dat$Age))
    
  keep <- c("Male", "Female", "Age")
  
  dat.sub <- dat[,keep]
  
  dat.melt <- melt(dat.sub, 
                   value.name='Population', 
                   variable.name = 'Gender', 
                   id.vars='Age' )
  
  h1 <- hPlot(
    y = 'Population', 
    x = 'Age', 
    type = 'bar', 
    data = dat.melt,
    group = 'Gender')
  
  h1$plotOptions(series = list(stacking = 'normal', pointPadding = 0, borderWidth = 0))
  
  h1$tooltip(formatter = "#! function() { return '<b>'+ this.series.name +', age '+ this.point.category +'</b><br/>' + 'Population: ' + Highcharts.numberFormat(Math.abs(this.point.y), 0);} !#")
  
  h1$legend(reversed = "true")
  
  if (max(dat.melt$Population >= 1000000)) {
    h1$yAxis(labels = list(formatter = "#! function() { return (Math.abs(this.value) / 1000000) + 'M';} !#"), 
             title = list(enabled = TRUE, text = 'Population'))
  } else {
    h1$yAxis(labels = list(formatter = "#! function() { return (Math.abs(this.value) / 1000) + 'K';} !#"), 
             title = list(enabled = TRUE, text = 'Population'))
  }
  
  if (!is.null(colors)) {
    h1$colors(colors)
  }
  if (length(year) > 1) {
    stop('Right now, hPyramid only accepts one year')
  }
  
  h1$exporting(enabled = TRUE)
  
  h1
}


# NVD3 pyramid

nPyramid <- function(country, year, colors = NULL) {
  dat <- getAgeTable(country, year)
  dat$Male <- -1 * dat$Male
  
  dat <- dat[order(rev(dat$ord)), ]
  
  keep <- c("Male", "Female", "Age")
  
  dat.sub <- dat[,keep]
  
  dat.melt <- melt(dat.sub, 
                   value.name='Population', 
                   variable.name = 'Gender', 
                   id.vars='Age' )
  
  dat.melt$abs <- abs(dat.melt$Population)
  
  n1 <- nPlot(
    y = 'Population', 
    x = 'Age', 
    group = 'Gender', 
    type = 'multiBarHorizontalChart', 
    data = dat.melt)
  
  # n1$xAxis(axisLabel = "Age") ## Need to work out label placement
  
  n1$chart(stacked = TRUE)
  
  n1$chart(tooltipContent = "#! function(key, x, y, e){
        var format = d3.format('0,000');
        return '<h3>' + key + ', age ' + x + '</h3>' + 
        '<p>' + 'Population: ' + format(e.point.abs) + '</p>'
        } !#")
  
    
  if (max(dat.melt$Population >= 1000000)) {    
    n1$yAxis(axisLabel = "Population",  
             tickFormat = "#! function(d) {
                          return d3.format(',.1f')(Math.abs(d) / 1000000) + 'M'
                          } !#")
  } else {
    n1$yAxis(axisLabel = "Population",  
             tickFormat = "#! function(d) {
                          return d3.format(',.0f')(Math.abs(d) / 1000) + 'K'
                          } !#")    
    
  }
  
  if (!is.null(colors)) {
    n1$chart(color = colors)
  }
  
  n1
}
```
